### 2019-07-06 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Done for improving Ch6

\begin{smallproof}
  Recall that we want to upper bound
  $ \ref{eq:5:term2} : = \sum_{k \in \Mbest} (\mu_k - \mu_{M*}) \left(T - \E_{\mu}[N_k(T)]\right)$.
  First, we observe that, for all $k\in \Mbest$,
  \begin{eqnarray*}
    T - \E_{\mu}[N_k(T)] & \leq T - \E_{\mu}\left[\sum_{t=1}^T \indic(\exists j : A^j(t) = k)\right] \\
    & = \E_{\mu}\left[\sum_{t=1}^T \indic(\forall j, A_j(t) \neq k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(k \notin \widehat{S}_t)\right],
  \end{eqnarray*}
  where we denote by $\widehat{S}_t = \{A^j(t), j \in [M]\}$ the set of selected arms at time $t$ (with no repetition). With this notation one can write
  \begin{eqnarray*}
  \ref{eq:5:term2} & \leq & (\mu_1 - \mu_{M^*})  \sum_{k \in \Mbest} \left(T - \E_{\mu}[N_k(T)]\right) \leq  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[\sum_{k \in \Mbest} \sum_{t = 1}^T \indic(k \notin \widehat{S}_t)\right] \\
  & = &  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right].
  \end{eqnarray*}
  The quantity $\sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)$ counts the number of optimal arms that have not been selected at time $t$. For each mis-selection of an optimal arm, there either exists a sub-optimal arm that has been selected, or an arm in $\Mbest$ on which a collision occurs. Hence
  \[\sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t) = \sum_{k \in \Mbest}\indic(C_k(t)) + \sum_{k \in \Mworst} \indic(\exists j : A^j(t) = k),\]
  which yields
  \[\E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right] \leq \sum_{k \in \Mbest}\E_{\mu}\left[\cC_k(T)\right] + \sum_{k \in \Mworst} \E_{\mu}\left[N_k(T)\right]\]
  and Lemma~\ref{lem:5:1stUpperBound} follows.
\end{smallproof}

@@

% \textbf{Index policies.}
% %
The approaches we now describe for multi-players bandits can be used in combination with any index policy (see Algorithm~\ref{algo:2:indexPolicy}), but we restrict our presentation to \UCB{} algorithms, for which strong theoretical guarantees can be obtained. In particular, we focus on two types of indices:
$\UCB_1$ \citep{Auer02}
and \klUCB{} \citep{KLUCBJournal}.
For player $j\in[M]$
%
denote $S_k^j(t) \eqdef \sum_{s=1}^t Y_{k,s} \indic(A^j(t) = k)$ the current sum of rewards obtained for arm $k$,
then $\widehat{\mu}_k^j(t) \eqdef S_k^j(t)/N_k^j(t)$ (if $N_k^j(t)\neq 0$) is the empirical mean of arm $k$, and thus one can define the index
\begin{equation}\label{eq:5:indexFor_UCB_klUCB}
  U_k^j(t) \eqdef \begin{cases}
      \widehat{\mu}_k^j(t)  + \sqrt{  f(t) / (2N_k^j(t))}
      &\text{for } \UCB, \\
      \sup\left\{ q \in [0, 1]: N_k^j(t)\times \kl(\widehat{\mu}_k^j(t), q) \leq f(t) \right\}
      &\text{for } \klUCB,
  \end{cases}
\end{equation}
where $f(t)$ is some \emph{exploration function}, usually taken to be $\log(t)$ in practice, and slightly larger in theory, which ensures that  $\Pr(U_k^j(t) \geq \mu_k) \gtrsim 1 - 1/t$ (see \cite{KLUCBJournal}).
A classical (single-player) UCB algorithm aims at the arm with largest index, as presented above in Section~\ref{sub:2:IndexPolicies}.
However, if each of the $M$ players selects the arm with largest UCB, all the players will end up colliding most of the time on the best arm.
To circumvent this problem, several coordination mechanisms have emerged, that rely on \emph{ordering} the indices and targeting \emph{one of} the $M$-best indices.

@@

We now focus on obtaining positive results for the algorithms we proposed in Section~\ref{sec:5:algorithms} above.
Section~\ref{sub:5:UpperBoundSelections} gives
an asymptotically optimal analysis of the expected number of sub-optimal draws
for our two proposals \RandTopM{} and \MCTopM{} as well as for \rhoRand{}, when they are combined with \klUCB{} indices,
and then we restrict to the most efficient algorithm \MCTopM, by proving in Section~\ref{sub:5:UpperBoundCollisions} that the number of collisions are logarithmic, and hence the regret of \MCTopM{} is also logarithmic.
%
Finally, Section~\ref{sub:5:SelfishFails} shortly discusses a disappointing result regarding \Selfish.
%  with more insights provided in Appendix~\ref{app:5:SelfishFails}.

@@

  \begin{lemma}\label{lem:5:Fact2KLUCB}
  \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
    For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $U_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
    \begin{small} % XXX
      \begin{align*}
        &\sum_{t=1}^T\Pr_{\bm{\mu}} \left( A^j(t) = k, \mu_{k'} \leq U_k^{j}(t) \right) & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
        + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.
      \end{align*}
    \end{small} % XXX
\end{leftbar}  % XXX leftbar lemmabar, comment if needed
\end{lemma}

@@

\begin{lemma}\label{lem:5:elementaryLemma_RandTopM_MCTopM}
\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
  For any $\boldsymbol{\mu}\in\cP_M$,
  any player $j \in [M]$ using
  \RandTopM- or \MCTopM-\klUCB,
  and any arm $k$,
  it holds that
  \begin{equation*}
    \sum_{t=1}^T
    \Pr\left(A^j(t)=k, k\notin \TopM(t)\right)
    = \left(\sum_{k', \mu_{k'} < \mu_k}\frac{1}{\kl(\mu_k,\mu_{k'})} + \sum_{k', \mu_{k'} > \mu_k}\frac{1}{\kl(\mu_{k'},\mu_{k})}\right) \log(T) + o(\log(T)).
  \end{equation*}
\end{leftbar}  % XXX leftbar lemmabar, comment if needed
\end{lemma}

@@

For the proposal \MCTopM, we are furthermore able to obtain a logarithmic regret upper bound, by proposing an original approach to control the number of collisions under this algorithm.
First, we can bound the number of collisions by the number of collisions for players not yet ``fixed on their arms'' ($\overline{s^j(t)}$),
that we can then bound by the number of changes of arms.

@@

Note that this bound is in $\bigO{M^3}$,
which significantly improves over the $\bigO{M{2M-1 \choose M}}$ bound, proven by \cite{Anandkumar11} for \rhoRand.
The obtained bound is worse than the $\bigO{M^2}$ proven by \cite{Rosenski16} for \MusicalChair{}. %, due to our trick of focusing on collisions for non-sitted players.
However, unlike \MusicalChair, our algorithm does not need any prior knowledge on the problem complexity.
The \MusicalChair{} algorithm is indeed not very satisfying from an applicative point of view, as it requires a prior knowledge of $\mu^*_{M}-\mu^*_{M+1}$.

@@

\paragraph{Logarithmic switching cost for \MCTopM.}\label{app:5:NumberSwitches}
%
While Lemma~\ref{lem:5:collisionsMCTopM} bounds the number of collisions,
another consequence of our proof is that we have also bounded
the (expected) number of \emph{switches of arms}, $\mathrm{SC}^{\cA}(T) = \sum_{t=1}^{T-1} \Pr(A^j(t+1) \neq A^j(t))$.
%
We controlled the total number of transitions $(2)$, $(3)$ and $(5)$ (see Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM}),
which are the only transitions when a player can switch from arm $k$ to arm $k'\neq k$.
Thus, the total number of arm switches is also proven to be logarithmic, if all players uses the \MCTopM-\klUCB{} algorithm.
This additional guarantee was never clearly stated for previous state-of-the-art works, like \rhoRand.
%
Even though minimizing the number of arms switching was not a goal,
this guarantee is appealing, in particular for Cognitive Radio applications,
where switching arms means re-configuring a radio hardware, an operation that costs energy.
An algorithm guaranteeing a small number of switches is thus interesting,
and for instance \cite{modiDemo2016} studied the empirical impact of the number of hardware reconfigurations on the battery life of a dynamic end-devices.
This work highlighted a tradeoff between minimizing regret and minimizing the number of switches of arms.
On a more experimental note, it was shown in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-players bandits, \rhoRand, could be tuned to run in batches, in order to reduce by a certain multiplicative factor its switching cost while multiplying its regret by the same factor.
While such ideas can interest a practitioner, they does not change the asymptotic behavior of $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, which is $\Theta(\log(T))$ for any efficient policy.
%
We also note that introducing explicit \emph{switching costs} in the regret, like it was done in previous works like \cite{Koren17} for single-player bandits, could also be fruitful.

@@

%
% Regular plots of centralized regrets
%
\begin{figure}[!t]
  % \centering
  % \begin{subfigure}[!h]{0.49\textwidth}
  %   \includegraphics[width=1.00\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_RegretCentralized____env1-1_3251433209347345969.pdf}
  % \end{subfigure}
  % % ~
  % \begin{subfigure}[!h]{0.49\textwidth}
    \includegraphics[width=0.95\textwidth]{MP__K9_M6_T5000_N500__4_algos/all_RegretCentralized____env1-1_8318947830261751207.pdf}
  % \end{subfigure}
  \caption[Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled.]{Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$. \textcolor{blue}{\rhoRand{} (top blue)} is outperformed by the other algorithms (and the gain increases with $M$). \textcolor{gold}{\MCTopM{} (bottom yellow)} outperforms all the other algorithms in most cases.}
  % \label{fig:5:MP__K9_M2-6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
  \label{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
  % \vspace*{-15pt}  % XXX remove if problem
\end{figure}

>>>>

\begin{smallproof}
  Recall that we want to upper bound
  $ \ref{eq:5:term2} : = \sum_{k \in \Mbest} (\mu_k - \mu_{M*}) \left(T - \E_{\mu}[N_k(T)]\right)$.
  First, we observe that, for all $k\in \Mbest$,
  % \begin{eqnarray*}
    $T - \E_{\mu}[N_k(T)] \leq T - \E_{\mu}\left[\sum_{t=1}^T \indic(\exists j : A^j(t) = k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(\forall j, A_j(t) \neq k)\right] = \E_{\mu}\left[\sum_{t=1}^T \indic(k \notin \widehat{S}_t)\right]$,
  % \end{eqnarray*}
  where we denote by $\widehat{S}_t = \{A^j(t), j \in [M]\}$ the set of selected arms at time $t$ (with no repetition). With this notation one can write
  \begin{eqnarray*}
  \ref{eq:5:term2} & \leq & (\mu_1 - \mu_{M^*})  \sum_{k \in \Mbest} \left(T - \E_{\mu}[N_k(T)]\right) \leq  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[\sum_{k \in \Mbest} \sum_{t = 1}^T \indic(k \notin \widehat{S}_t)\right] \\
  & = &  (\mu_1 - \mu_{M^*})  \E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right].
  \end{eqnarray*}
  The quantity $\sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)$ counts the number of optimal arms that have not been selected at time $t$. For each mis-selection of an optimal arm, there either exists a sub-optimal arm that has been selected, or an arm in $\Mbest$ on which a collision occurs. Hence
  \[\sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t) = \sum_{k \in \Mbest}\indic(C_k(t)) + \sum_{k \in \Mworst} \indic(\exists j : A^j(t) = k),\]
  which yields
  \[\E_{\mu}\left[ \sum_{t = 1}^T \sum_{k \in \Mbest}\indic(k \notin \widehat{S}_t)\right] \leq \sum_{k \in \Mbest}\E_{\mu}\left[\cC_k(T)\right] + \sum_{k \in \Mworst} \E_{\mu}\left[N_k(T)\right]\]
  and Lemma~\ref{lem:5:1stUpperBound} follows.
\end{smallproof}

@@

% \textbf{Index policies.}
% %
The approaches we now describe for multi-players bandits can be used in combination with any index policy (see Algorithm~\ref{algo:2:indexPolicy}), but we restrict our presentation to \UCB{} algorithms, for which strong theoretical guarantees can be obtained. In particular, we focus on two types of indices:
$\UCB_1$ \citep{Auer02}
and \klUCB{} \citep{KLUCBJournal}.
For player $j\in[M]$
%
denote $S_k^j(t) \eqdef \sum_{s=1}^t Y_{k,s} \indic(A^j(t) = k)$ the current sum of rewards obtained for arm $k$,
then $\widehat{\mu}_k^j(t) \eqdef S_k^j(t)/N_k^j(t)$ (if $N_k^j(t)\neq 0$) is the empirical mean of arm $k$, and thus one can define the index
\begin{equation}\label{eq:5:indexFor_UCB_klUCB}
  U_k^j(t) \eqdef \begin{cases}
      \widehat{\mu}_k^j(t)  + \sqrt{  f(t) / (2N_k^j(t))}
      &\text{for } \UCB, \\
      \sup\left\{ q \in [0, 1]: N_k^j(t)\times \kl(\widehat{\mu}_k^j(t), q) \leq f(t) \right\}
      &\text{for } \klUCB,
  \end{cases}
\end{equation}
where $f(t)$ is some \emph{exploration function}, usually taken to be $\log(t)$ in practice, and slightly larger in theory (\eg, $f(t)=\log(t)+3\log(\log(t))$), which ensures that  $\Pr(U_k^j(t) \geq \mu_k) \gtrsim 1 - 1/t$ (see \cite{KLUCBJournal}).
A classical (single-player) UCB algorithm aims at the arm with largest index, as presented above in Section~\ref{sub:2:IndexPolicies}.
However, if each of the $M$ players selects the arm with largest UCB, all the players will end up colliding most of the time on the best arm.
To circumvent this problem, several coordination mechanisms have emerged, that rely on \emph{ordering} the indices and targeting \emph{one of} the $M$-best indices.

@@

We now focus on obtaining positive results for the algorithms we proposed in Section~\ref{sec:5:algorithms} above.
Section~\ref{sub:5:UpperBoundSelections} gives
an asymptotically optimal analysis of the expected number of sub-optimal draws
for our two proposals \RandTopM{} and \MCTopM{} as well as for \rhoRand{}, when they are combined with \klUCB{} indices,
and then we restrict to the most efficient algorithm \MCTopM, by proving in Section~\ref{sub:5:UpperBoundCollisions} that it achieves a logarithmic number of collisions and regret.
%
Finally, Section~\ref{sub:5:SelfishFails} shortly discusses a negative result regarding \Selfish.
%  with more insights provided in Appendix~\ref{app:5:SelfishFails}.

@@

  \begin{lemma}\label{lem:5:Fact2KLUCB}
  \begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
    For any arms $k$ and $k'$ such that $\mu_{k'} > \mu_{k}$, if $U_k^{j}(t)$ is the kl-UCB index with exploration function $f(t)$,
    % \begin{small} % XXX
      \begin{align}
        \sum_{t=1}^T\Pr_{\bm{\mu}} \left( A^j(t) = k, \mu_{k'} \leq U_k^{j}(t) \right)
        & \leq  \frac{f(T)}{\kl(\mu_k,\mu_{k'})}
        + \sqrt{2\pi} \sqrt{\frac{\kl'(\mu_k,\mu_{k'})^2}{\kl(\mu_k,\mu_{k'})^3}}\sqrt{f(T)} \nonumber\\
        & + 2\left(\frac{\kl'(\mu_k,\mu_{k'})}{\kl(\mu_k,\mu_{k'})}\right)^2 + 1.
      \end{align}
    % \end{small} % XXX
\end{leftbar}  % XXX leftbar lemmabar, comment if needed
\end{lemma}

@@

\begin{lemma}\label{lem:5:elementaryLemma_RandTopM_MCTopM}
\begin{leftbar}[lemmabar]  % XXX leftbar lemmabar, comment if needed
  For any $\boldsymbol{\mu}\in\cP_M$,
  any player $j \in [M]$ using
  \RandTopM- or \MCTopM-\klUCB,
  and any arm $k$,
  it holds that
  \begin{align}
    \sum_{t=1}^T \Pr\left(A^j(t)=k, k\notin \TopM(t)\right) =
    & \left( \sum_{k', \mu_{k'} < \mu_k}\frac{1}{\kl(\mu_k,\mu_{k'})} \right) \log(T) \nonumber\\
    & \left( \sum_{k', \mu_{k'} > \mu_k}\frac{1}{\kl(\mu_{k'},\mu_{k})}\right ) \log(T)
    + o(\log(T)).
  \end{align}
\end{leftbar}  % XXX leftbar lemmabar, comment if needed
\end{lemma}

@@

For \MCTopM, we are furthermore able to obtain a logarithmic regret upper bound, by proposing an original approach to control the number of collisions under this algorithm.
First, we can bound the number of collisions by the number of collisions for players not yet ``fixed on their arms'' ($\overline{s^j(t)}$),
that we can then bound by the number of changes of arms.

@@

Note that this bound is in $\bigO{M^3}$,
which significantly improves over the $\bigO{M{2M-1 \choose M}}$ bound, proven by \cite{Anandkumar11} for \rhoRand.
The obtained bound is worse than the $\bigO{M^2}$ proven by \cite{Rosenski16} for \MusicalChair{}. %, due to our trick of focusing on collisions for non-sitted players.
However, unlike \MusicalChair, our algorithm does not need any prior knowledge on the problem complexity,
at it is indeed not very satisfying from an applicative point of view to require a prior knowledge of $\mu^*_{M}-\mu^*_{M+1}$.

@@

\paragraph{Logarithmic switching cost for \MCTopM.}\label{app:5:NumberSwitches}
%
While Lemma~\ref{lem:5:collisionsMCTopM} bounds the number of collisions,
another consequence of our proof is that we have also bounded
the (expected) number of \emph{switches of arms}, $\mathrm{SC}^{\cA}(T) = \sum_{t=1}^{T-1} \Pr(A^j(t+1) \neq A^j(t))$.
%
We controlled the total number of transitions $(2)$, $(3)$ and $(5)$ (see Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM}),
which are the only transitions when a player can change its arm.
%  switch from arm $k$ to arm $k'\neq k$.
Thus, the total number of arm switches is also logarithmic, if the $M$ players use \MCTopM-\klUCB.
This additional guarantee was never clearly stated for previous works, like \rhoRand.
%
Even though minimizing the number of arms switching was not a goal,
this guarantee is appealing, in particular for Cognitive Radio applications,
where switching arms means re-configuring a radio hardware, an operation that costs energy.
An algorithm guaranteeing a small number of switches is thus interesting,
and for instance \cite{modiDemo2016} studied the empirical impact of the number of hardware reconfigurations on the battery life of a dynamic end-devices.
This work highlighted a tradeoff between minimizing regret and minimizing the number of switches of arms.
On a more experimental note, it was shown in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-players bandits, \rhoRand, could be tuned to run in batches, in order to reduce by a certain multiplicative factor its switching cost while multiplying its regret by the same factor.
While such ideas can interest a practitioner, they does not change the asymptotic behavior of $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, which is $\Theta(\log(T))$ for any efficient policy.
%
We also note that introducing explicit \emph{switching costs} in the regret, like it was done in previous works like \cite{Koren17} for single-player bandits, could also be fruitful.

@@

%
% Regular plots of centralized regrets
%
\begin{figure}[!t]
  % \centering
  % \begin{subfigure}[!h]{0.49\textwidth}
  %   \includegraphics[width=1.00\textwidth]{MP__K9_M2_T5000_N500__4_algos/all_RegretCentralized____env1-1_3251433209347345969.pdf}
  % \end{subfigure}
  % % ~
  % \begin{subfigure}[!h]{0.49\textwidth}
    \includegraphics[width=1.05\textwidth]{MP__K9_M6_T5000_N500__4_algos/all_RegretCentralized____env1-1_8318947830261751207.pdf}
  % \end{subfigure}
  \caption[Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled.]{Regret for $M=6$ players, $K=9$ arms, horizon $T=5000$, against $500$ problems $\boldsymbol{\mu}$ uniformly sampled in $[0,1]^K$. \textcolor{blue}{\rhoRand{} (top blue)} is outperformed by the other algorithms (and the gain increases with $M$). \textcolor{gold}{\MCTopM{} (bottom yellow)} outperforms all the other algorithms in most cases.}
  % \label{fig:5:MP__K9_M2-6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
  \label{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems}
  % \vspace*{-15pt}  % XXX remove if problem
\end{figure}
